---
title: "Reproducible Research Project"
author: "Bruno Martins"
date: "24/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

This is the Bruno's project for Reproducible Reserach Course, on Coursera.

### Reading/processing the dataset

```{r echo=T, message=FALSE}

library(dplyr)
library(ggplot2)

act <- read.csv("activity.csv")
head(act)

```

Below the dataset wiht steps consolidated by day:

```{r echo=T}

actDay <- act %>%
  group_by(date) %>%
  summarise(steps_by_day = sum(steps, na.rm = T), mean_steps_by_day = mean(steps, na.rm = T), 
            median_steps_by_day = median(steps, na.rm = T)) %>%
  mutate(day = as.Date(date)) %>%
  ungroup()

head(actDay)

```

### Histogram of the total number of steps taken each day

```{r echo=T, fig.cap = "Fig 1:  Histogram of the total number of steps taken each day"}

library(ggplot2)
qplot(data = actDay,steps_by_day, bins = 20)

```
### Mean and median number of steps taken each day

Mean: 
```{r echo=T, fig.cap = "Fig 1:  Mean number of steps taken each day"}

q <- ggplot(data = actDay, aes(x = day, y = mean_steps_by_day)) +
  geom_bar(stat="identity")
q

```

Median: 
```{r echo=T, fig.cap = "Fig 1:  Median number of steps taken each day"}

q <- ggplot(data = actDay, aes(x = day, y = median_steps_by_day)) +
 geom_line(aes(group = 1))
q

```

### Time series plot of the average number of steps taken

```{r echo=T, fig.cap = "Fig 1:  Mean and median number of steps taken each day"}

q <- ggplot(data = actDay, aes(x = day, y = mean_steps_by_day)) +
  geom_line(aes(group = 1))
q

```

### The 5-minute interval that, on average, contains the maximum number of steps

The 835 interval has, on average, 206.1698,and is the one which has the maximum number of steps.

```{r echo=T}

intStp <- act %>%
  group_by(interval) %>%
  select(-date) %>%
  summarise(steps = mean(steps, na.rm = T)) %>%
  arrange(desc(steps))

intStp[is.na(intStp)] <- 0
head(intStp)

```
### Code to describe and show a strategy for imputing missing data


I used the "is.na" function to replace the NA's for 0. Below a exemplo.

```{r, echo=T}
act2 <- act
NAsBefore <- is.na(act2$steps)
table(NAsBefore)
```

The original content has 2304 missin values (NAs). Below the result after the replacement.

```{r, echo=T}

act2[is.na(act2)] <- 0
NAsAfter <- is.na(act2$steps)
table(NAsAfter)

```
### Histogram of the total number of steps taken each day after missing values are imputed

```{r echo=T, fig.cap = "Fig 1:  Histogram of the total number of steps taken each day after missing values are imputed"}


actDay[is.na(actDay)] <- 0

library(ggplot2)
qplot(data = actDay,steps_by_day, bins = 20)

```

### Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekend

```{r, echo = T}

actDay$weekday <- weekdays(actDay$day)
head(actDay, 10)

weekend <- c("sábado", "domingo")
inWnd <- actDay$weekday  %in% weekend

actDayWnd <- actDay[inWnd,]
actDayW <- actDay[-inWnd,]

par(2,1)

q1 <- ggplot(data = actDayW, aes(x = day, y = mean_steps_by_day)) +
  geom_line(aes(group = 1)) + ggtitle("Average steps on weekdays")

q2 <- ggplot(data = actDayWnd, aes(x = day, y = mean_steps_by_day)) +
  geom_line(aes(group = 1)) + ggtitle("Average steps on weekends")

require(gridExtra)
grid.arrange(q1, q2, ncol=2)

```


